---
title: "Crime Notebook"
output: html_notebook
---
OBJECTIVE:
AN ANALYTICAL DETECTIVE
Crime is an international concern, but it is documented and handled in very different ways in different countries. In the United States, violent crimes and property crimes are recorded by the Federal Bureau of Investigation (FBI). Additionally, each city documents crime, and some cities release data regarding crime rates. The city of Chicago, Illinois releases crime data from 2001 onward online.

What is the local language?
```{r}
Sys.getlocale()
```
Change the local to English
```{r}
Sys.setlocale("LC_ALL","English")
```


Useful libraries

```{r}
library("dplyr")
library("ggplot2")
```

Read the dataset mvtWeek1.csv into R, using the read.csv function, and call the data frame "mvt". Remember to navigate to the directory on your computer containing the file mvtWeek1.csv first. It may take a few minutes to read in the data, since it is pretty large. Then, use the str and summary functions to answer the following questions

```{r}
# import data
data <- read.csv("../data/mvtWeek1.csv")

# create data frame
mvt <- tbl_df(data)

# Analyse data
head(mvt)
```

LOADING THE DATA 
How many rows of data (observations) are in this dataset?

```{r}
N = nrow(mvt)
L = ncol(mvt)
str(df)
```
Using the "max" function, what is the maximum value of the variable "ID"?

```{r}
max(mvt$ID)
```

What is the minimum value of the variable "Beat"?

```{r}
min(mvt$Beat)
```

Simplification

```{r}
 summary(mvt)
```

How many observations have a LocationDescription value of ALLEY?
```{r}
summary(mvt$LocationDescription)
```

In many datasets, like this one, you have a date field. Unfortunately, R does not automatically recognize entries that look like dates. We need to use a function in R to extract the date and time. Take a look at the first entry of Date (remember to use square brackets when looking at a certain entry of a variable).

In what format are the entries in the variable Date?
```{r}
head(mvt)
```

first line: 12/31/12 23:15
Answer -> Month/Day/Year Hour:Minute

let's convert these characters into a Date object in R

```{r}
DateConvert =  as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))
```

This converts the variable "Date" into a Date object in R. Take a look at the variable DateConvert using the summary function.
What is the month and year of the median date in our dataset? Enter your answer as "Month Year", without the quotes. (Ex: if the answer was 2008-03-28, you would give the answer "March 2008", without the quotes.)


```{r}
summary(DateConvert)
```

2006-05-21 Gives May 2006
UNDERSTANDING DATES IN R 
let's extract the month and the day of the week, and add these variables to our data frame 


```{r}
mvt$Month = months(DateConvert)
mvt$Weekday = weekdays(DateConvert)
```

This creates two new variables in our data frame, Month and Weekday, and sets them equal to the month and weekday values that we can extract from the Date object. Lastly, replace the old Date variable with DateConvert by typing:

```{r}
mvt$Date = DateConvert
```

Using the table command, answer the following questions.
In which month did the fewest motor vehicle thefts occur?

```{r}
sort(table(mvt$Month))
```

On which weekday did the most motor vehicle thefts occur?

```{r}
sort(table(mvt$Weekday))
```

Each observation in the dataset represents a motor vehicle theft, and the Arrest variable indicates whether an arrest was later made for this theft. 

Which month has the largest number of motor vehicle thefts for which an arrest was made?

Method 1:

```{r}
table(mvt$Month,mvt$Arrest)
```

Method 2:
Group by month, by arrest
TBD

```{r}
by_MonthArrest <- group_by(mvt,Month,Arrest)
by_MonthArrest %>%  summarise(N = n()) %>% 
  select(Month,Arrest,N) %>%
  filter(Arrest == TRUE) -> df_N
df_N
```

VISUALIZING CRIME TRENDS 

Now, let's make some plots to help us better understand how crime has changed over time in Chicago. Throughout this problem, and in general, you can save your plot to a file. For more information, this website very clearly explains the process.
First, let's make a histogram of the variable Date. We'll add an extra argument, to specify the number of bars we want in our histogram. In your R console, type

```{r}
hist(mvt$Date,breaks = 100,xlab = "Date")
```

Analysis:
In general, does it look like crime increases or decreases from 2002 - 2012? DECREASES
In general, does it look like crime increases or decreases from 2005 - 2008? INCREASES


Now, let's see how arrests have changed over time. Create a boxplot of the variable "Date", sorted by the variable "Arrest" (if you are not familiar with boxplots and would like to learn more, check out this tutorial). In a boxplot, the bold horizontal line is the median value of the data, the box shows the range of values between the first quartile and third quartile, and the whiskers (the dotted lines extending outside the box) show the minimum and maximum values, excluding any outliers

(which are plotted as circles). Outliers are defined by first computing the difference between the first and third quartile values, or the height of the box. This number is called the Inter-Quartile Range (IQR). Any point that is greater than the third quartile plus the IQR or less than the first quartile minus the IQR is considered an outlier.
Does it look like there were more crimes for which arrests were made in the first half of the time period or the second half of the time period? (Note that the time period is from 2001 to 2012, so the middle of the time period is the beginning of 2007.)


