---
title: "Linear Random"
output:
  html_document: default
  html_notebook: default
---

## Context
This is probably the dumbest dataset on Kaggle. The whole point is, however, to provide a common dataset for linear regression. Although such a dataset can easily be generated in Excel with random numbers, results would not be comparable.

## Content
The training dataset is a CSV file with 700 data pairs (x,y). The x-values are numbers between 0 and 100. The corresponding y-values have been generated using the Excel function NORMINV(RAND(), x, 3). Consequently, the best estimate for y should be x. The test dataset is a CSV file with 300 data pairs.

## Load

```{r}
library(dplyr)
library(ggplot2)
datatrain = read.csv("../data/train.csv")
train = tbl_df(datatrain)
datatest = read.csv("../data/test.csv")
test = tbl_df(datatest)
```

## Analysis training
```{r}
train <- na.omit(train)
str(train)
summary(train)
cor(train)
```

## Visualisation

```{r}
plot(x = train$x,y = train$y,type = "p",col="red",main="Train y vs x")
```

## Analysis test
```{r}
test <- na.omit(test)
str(test)
summary(test)
cor(test)
```

## Visualisation

```{r}
plot(x = test$x,y = test$y,type = "p",col="blue",main="Test y vs x")
```

```{r Linear Model}
model = lm(y ~ x,data = train)
summary(model)
```

```{r Predictions}
predTest = predict(model, newdata=test)
summary(predTest)
```

What is the sum of squared errors (SSE) of lmScore on the testing set?

```{r}
SSE = sum((predTest-test$y)^2)
```

What is the root-mean squared error (RMSE) of lmScore on the testing set?

```{r}
RMSE = sqrt(mean((predTest-test$y)^2))
```

What is the predicted test score used in the baseline model? 
```{r}
baseline = mean(test$y)
```

What is the sum of squared errors of the baseline model on the testing set?
```{r}
SST = sum((baseline-test$y)^2)
```

What is the test-set R-squared value of model?

```{r}
testSetR2 =  1-SSE/SST
testSetR = sqrt(testSetR2)
```

